### 2018-01-22 acmart-master/sample-sigconf-authordraft.tex
Started Use Case

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{tikz}
\usepackage{pgf-umlsd}
\usepgflibrary{arrows}
\usetikzlibrary{calc}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{}

% ISBN
\acmISBN{}

%Conference
\acmConference[e-Energy'18]{ACM e-Energy conference}{July 2018}{Karlsruhe, Germany}
\acmYear{2018}
\copyrightyear{2018}

\acmArticle{4}
\acmPrice{15.00}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}
%\editor{Theo D'Hondt}
%\editor{Wolfgang De Meuter}

\begin{document}
\title{Usable Security in Virtual Power Plants}
%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as  \texttt{acmart.pdf} document}

\author{Amit Dilip Patil}
%\authornote{Dr.~Trovato insisted his name be first.}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{University of Passau}
  %\streetaddress{P.O. Box 1212}
  \city{Passau}
  \state{Germany}
  %\postcode{43017-6221}
}
\email{Amit.Patil@uni-passau.de}

\author{Hermann De Meer}
%\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
  \institution{University of Passau}
  \streetaddress{P.O. Box 1212}
  \city{Passau}
  \state{Germany}
  %\postcode{43017-6221}
}
\email{	Hermann.DeMeer@uni-passau.de}

% The default list of authors is too long for headers.
\renewcommand{\shortauthors}{Patil et al.}

\begin{abstract}
Automation of systems requires users to be less involved in system control. During situations when the user must interface with the system by themselves, the technology available must assist the user to complete their task and must not be a hindrance. Therefore, it is necessary to make the system more usable. Usability is often overlooked in favour of primary system goals, hence the lack of advancement in the usability drive. There exists no concrete framework to assess usability of a system, since most existing approaches focus on user feedback and neglect system optimization. In this study, usability is described in context of security in virtual power plants. A usability framework was established for authentication mechanisms. It was determined that usability arises by optimizing system factors that relate to usability standards with the crucial addition of user preferences. Additionally, the approach can be applied to any target system to assess usability.
\end{abstract}

% \footnote{This is an abstract footnote}
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002978.10003029.10011703</concept_id>
<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10002978.10002991.10002992</concept_id>
<concept_desc>Security and privacy~Authentication</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10003120.10003123.10011760</concept_id>
<concept_desc>Human-centered computing~Systems and tools for interaction design</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003120.10003121.10003122.10010854</concept_id>
<concept_desc>Human-centered computing~Usability testing</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Security and privacy~Usability in security and privacy}
\ccsdesc[300]{Security and privacy~Authentication}
\ccsdesc[500]{Human-centered computing~Systems and tools for interaction design}
\ccsdesc[300]{Human-centered computing~Usability testing}

\keywords{Security, Usability, Virtual Power Plants}

\maketitle

%\input{samplebody-conf}
\input{draft}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography}

\end{document}