### 2018-01-25 acmart-master/sample-sigconf-authordraft.tex

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{tikz}
\usepackage{pgf-umlsd}
\usepgflibrary{arrows}
\usetikzlibrary{calc}
\usepackage{amsmath}
\usepackage{amssymb}
% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{}

% ISBN
\acmISBN{}

%Conference
\acmConference[e-Energy'18]{ACM e-Energy conference}{July 2018}{Karlsruhe, Germany}
\acmYear{2018}
\copyrightyear{2018}

\acmArticle{4}
\acmPrice{15.00}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}
%\editor{Theo D'Hondt}
%\editor{Wolfgang De Meuter}

\begin{document}
\title{Usable Security in Virtual Power Plants}
%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as  \texttt{acmart.pdf} document}

\author{Amit Dilip Patil}
%\authornote{Dr.~Trovato insisted his name be first.}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{University of Passau}
  %\streetaddress{P.O. Box 1212}
  \city{Passau}
  \state{Germany}
  %\postcode{43017-6221}
}
\email{Amit.Patil@uni-passau.de}

\author{Hermann De Meer}
%\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
  \institution{University of Passau}
  \streetaddress{P.O. Box 1212}
  \city{Passau}
  \state{Germany}
  %\postcode{43017-6221}
}
\email{	Hermann.DeMeer@uni-passau.de}

% The default list of authors is too long for headers.
\renewcommand{\shortauthors}{Patil et al.}

\begin{abstract}
Power grid systems deliver electricity from suppliers that generate power to customers
including private homes, corporations and public domain. The transmission takes place
through electric grids and networks that span across geographical areas. Smart grids utilize information and communication technology to perform the required functions where
SCADA(Supervisory Control And Data Acquisition) systems monitor, control and manage
the processes in critical infrastructures. They include Virtual Power Plants(VPPs) that
schedule energy sources appropriately to meet the energy demand. Dynamic Virtual Power
Plants(DVPPs) have been proposed to decentralize VPPs to make energy scheduling more
flexible. As with any critical infrastructure, attacks on VPPs and DVPPs could result in
catastrophic situations. Security is required in place to defend the system against attacks. However, security mechanisms such as authentication systems may actually prevent the operator from gaining access to the system in time to control the power plant. In such a case, security is actually a hindrance. The authentication mechanism must be more usable from the perspective of the operator, even under stressful situations. Therefore, the aim of this paper is to establish a method to approach usable security by design.
\end{abstract}

% \footnote{This is an abstract footnote}
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002978.10003029.10011703</concept_id>
<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10002978.10002991.10002992</concept_id>
<concept_desc>Security and privacy~Authentication</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10003120.10003123.10011760</concept_id>
<concept_desc>Human-centered computing~Systems and tools for interaction design</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003120.10003121.10003122.10010854</concept_id>
<concept_desc>Human-centered computing~Usability testing</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Security and privacy~Usability in security and privacy}
\ccsdesc[300]{Security and privacy~Authentication}
\ccsdesc[500]{Human-centered computing~Systems and tools for interaction design}
\ccsdesc[300]{Human-centered computing~Usability testing}

\keywords{Security, Usability, Virtual Power Plants}

\maketitle

%\input{samplebody-conf}
\input{draft}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography}

\end{document}